---
title: "p8105_hw2_rac2224"
author: "Rachel Crowley"
date: "10/3/2019"
output: github_document
---

## Problem 1

Reading and cleaning the Mr. Trash Wheel dataset

```{r}
library(tidyverse)
library(readxl)
library(knitr)
trash_data = read_excel("./data/Trashwheel.xlsx", sheet = 1, range = "A2:N408",   col_types =   NULL) 
  trash_data = janitor::clean_names(trash_data) 
  trash_data = drop_na(trash_data, dumpster) 
  trash_data = mutate(trash_data, 
                      sports_balls = round(sports_balls, digits = 0)) 
  as.integer(pull(trash_data, sports_balls))
```

Reading and cleaning precipitation datasets for 2017 and 2018

```{r}
prec2018_data = read_excel("./data/Trashwheel.xlsx", sheet = 5, range =           "A2:B14", col_types = NULL)
  prec2018_data = janitor::clean_names(prec2018_data)
  prec2018_data = mutate(prec2018_data,
                         year = "2018")
prec2017_data = read_excel("./data/Trashwheel.xlsx", sheet = 6, range =           "A2:B14", col_types = NULL)
  prec2017_data = janitor::clean_names(prec2017_data)
  prec2017_data = mutate(prec2017_data,
                         year = "2017")
```

Combining precipitation datasets and converting month to character variable

```{r}
precip_data = 
  full_join(prec2017_data, prec2018_data)
precip_data = transform(precip_data, month = month.name)
```

* The number of observations in the dataset precip_data is `r nrow(precip_data)`. The number of observations in the dataset trash_data is `r nrow(trash_data)`.Within the dataset precip_data, key variables include `r ls(precip_data)`. Within the dataset trash_data, key variables include `r ls(trash_data)`.The total precipitation in 2018 is `r sum(prec2018_data$total)`. The median number of sports balls in a dumpster in 2017 is `r  median(subset(trash_data, year==2017)$sports_balls)`. 

## Problem 2

Cleaning data in FiveThirtyEight

Cleaning pols-month dataset

```{r}
polsmonth_data=read_csv(file = "./data/pols-month.csv")
polsmonth_data = janitor::clean_names(polsmonth_data)
polsmonth_data = separate(polsmonth_data, 
  mon, into=c("year", "month","day"))
polsmonth_data = mutate(polsmonth_data, month=month.name[as.integer(month)]) 
polsmonth_data = mutate(polsmonth_data, president = prez_gop + prez_dem) %>%
  select( -prez_gop, -prez_dem, -day)  
```

Cleaning snp dataset

```{r}
snp_data=read_csv(file = "./data/snp.csv")
snp_data = janitor::clean_names(snp_data)
snp_data = separate(snp_data, 
  date, into=c("month", "day","year"))
snp_data = mutate(snp_data, month=month.name[as.integer(month)]) %>%
  select(-day) 
snp_data = select(snp_data, year, month, everything())
snp_data = head(arrange(snp_data, year, month), 787)
```

Cleaning unemployment dataset

```{r}
unemployment_data=read_csv(file = "./data/unemployment.csv")
unemployment_data = janitor::clean_names(unemployment_data)
unemployment_data = 
    pivot_longer(
    unemployment_data,
    jan:dec,
    names_to = "month",
    values_to = "unemployment")
unemployment_data = transform(unemployment_data, month = month.name)
```
Merging snp, employment, and pols datasets

```{r}
unemployment_data <- unemployment_data %>%  
  mutate(year = as.character(year))
polsnp_data =
  left_join(polsmonth_data, snp_data)
merged_data = 
  left_join(polsnp_data, unemployment_data)
```


* The polsmonth dataset has the following range of years: a minimum year of `r min(pull(polsmonth_data, year))` and a maximum year of `r max(pull(polsmonth_data, year))`. The dimensions of the polsmonth dataset are `r dim(polsmonth_data)`. In terms of key variables, within the dataset polsmonth the following key variables were present: `r ls(polsmonth_data)`. 

* The snp dataset has the following range of years: a minimum year of `r min(pull(snp_data, year))` and a maximum year of `r max(pull(snp_data, year))`. The dimensions of the snp dataset are `r dim(snp_data)`. The following key variables were present within the snp dataset: `r ls(snp_data)`. 

* The unemployment dataset has the following range of years: a minimum year of `r min(pull(unemployment_data, year))` and a maximum year of `r max(pull(unemployment_data, year))`.The dimensions of the unemployment dataset are `r dim(unemployment_data)`.
The following key variables were present within the unemployment dataset: `r ls(unemployment_data)`. 

* The resulting dataset, merged_data, has the following range of years: a minimum year of `r min(pull(merged_data, year))` to a maximum year of `r max(pull(merged_data,year))`. The dimensions of the resulting dataset, merged_data, are `r dim(merged_data)`. The following key variables were present within the resulting dataset, merged_data: `r ls(merged_data)`.


## Problem 3

Cleaning Popular Baby Names dataset

```{r}
babynames_data=read_csv(file = "./data/Popular_Baby_Names.csv")
babynames_data = janitor::clean_names(babynames_data)
babynames_data = babynames_data %>% distinct()
babynames_data$ethnicity <- gsub("BLACK NON HISPANIC", "BLACK NON HISP", babynames_data$ethnicity)
babynames_data$ethnicity <- gsub("WHITE NON HISPANIC", "WHITE NON HISP", babynames_data$ethnicity)
babynames_data$ethnicity <- gsub("ASIAN AND PACIFIC ISLANDER", "ASIAN AND PACI", babynames_data$ethnicity)
babynames_data= mutate(babynames_data, childs_first_name = str_to_title(childs_first_name))
```

Table for showing rank of popularity in name "Olivia" over time 

```{r}
olivia_data =  babynames_data[which(babynames_data$childs_first_name=='Olivia'), 1:6]
olivia_data = head(arrange(olivia_data, year_of_birth), 24) %>%
  select(-gender, -count, -childs_first_name) 
olivia_data = 
  pivot_wider(
  olivia_data, 
  names_from = "year_of_birth", 
  values_from = "rank")
kable(olivia_data[1:4, ], format = "pandoc", caption = "Rank in popularity of name Olivia over time")
```

Table for most popular name among male children over time

```{r}
male_data = babynames_data[which(babynames_data$gender=='MALE' & babynames_data$rank=="1"), 1:6]
male_data = head(arrange(male_data, year_of_birth), 5962) %>%
    select(-gender, -count, -rank) 
male_data = 
  pivot_wider(
male_data, 
  names_from = "year_of_birth", 
  values_from = "childs_first_name")
kable(male_data[1:4,], format = "pandoc", caption = "Most popular name among male children over time")
```

Scatter plot for male, white non-hispanic children born in 2016 

```{r}
whitemale_data= babynames_data[which(babynames_data$gender=='MALE' & babynames_data$year_of_birth=="2016" & babynames_data$ethnicity=="WHITE NON HISP"), 1:6]
whitemale_plot = ggplot(whitemale_data, aes(x = rank, y = count)) +
  geom_point() + ggtitle("Name rank vs. count among male, white non-hispanic children born in 2016")
print(whitemale_plot)
ggsave("whitemale_plot.png")
```